---
output:
  tufte::tufte_handout: default
---
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, fig.width = 5, fig.height = 3.75)
```

```{r wrangling_cruise2, include=FALSE}
property <- "pierce"
year <- 2019

load(file = paste("C:/Users/Neal/projects/cruise/rda/", property, "-cruise-", year, ".rda", sep = ""))

library(tidyverse)
library(ggthemes)
library(stargazer)
library(knitr)
library(kableExtra)
library(extrafont)
loadfonts("win", quiet = TRUE)
```

```{r set1}
std <- 1
```

#Area `r std` 
`r (stands %>% filter(stand == std))$type`    
\noindent `r (stands %>% filter(stand == std))$acres_legal` legal acres | `r (stands %>% filter(stand == std))$acres_calc` measured acres    

##Site-specific information

> - __Soils:__   
\indent\indent `r paste((soils %>% filter(stand == std))$text, collapse = " ") %>% str_replace_all("START", "\n") %>% str_replace_all("END", "  ")` 

> - __Site Class:__   
\vspace{2pt}
`r (stands %>% filter(stand == std))$site_class` (determined from soil mapping and field assessment)   

> - __Access:__   
\vspace{2pt}
`r (stands %>% filter(stand == std))$access` 

> - __Stand history:__   
\vspace{2pt}
`r (stands %>% filter(stand == std))$history` 

##Current forest information

> - __Age Class Structure:__   
\vspace{2pt}
`r (stands %>% filter(stand == std))$structure.1`  
```{r diam_dist1, fig.margin=TRUE, fig.width=2, fig.height=6, fig.cap="Distributions are approximated with kernel density estimation. Common species are those that account for at least 8 percent of the total stocking and areas under each curve represent species basal areas."}

# diameter dist of main spp in a specific stand ----

trees %>% filter(stand == std, live == 1, dbh >= 6) %>%
  group_by(spp) %>%
  mutate(spp_sum = sum(live)) %>%
  ungroup() %>%
  mutate(spp = reorder(spp, -live, FUN = sum)) %>%
  filter(spp_sum/n()>=.08) %>% 
  ggplot(aes(dbh, y = ..count..)) +
  geom_density(alpha = .6, position = "identity", fill = "grey", col = "dark grey") +
  theme(text = element_text(family = "Perpetua"), 
        legend.position = "none") +
  facet_wrap(~spp, ncol = 1) +
  scale_y_continuous("relative basal area", labels = NULL, 
                     breaks = NULL) +
  scale_x_continuous("dbh") +
  labs(title = "Diameter distributions", 
       subtitle = "for common species")
```

```{r spp}
x <- trees %>% 
  filter(stand == std, live == 1, dbh >= 6) %>% 
  group_by(spp) %>% 
  summarize(spp_sum = sum(live)) %>%
  mutate(pct = spp_sum / sum(spp_sum)) %>%
  select(spp, pct) %>%
  arrange(-pct)
```

> - __Species (% stocking):__   
\vspace{2pt}
`r if_else(nrow(x)>0, paste(x$spp, " (", round(x$pct*100), '%)', sep = '', collapse = ", "), "none")`  

> - __Regeneration:__   
\vspace{2pt}
`r (stands %>% filter(stand == std))$regen`  

> - __Forest health:__   
\vspace{2pt}
`r (stands %>% filter(stand == std))$health`  

> - __Volume/ac:__   
\vspace{2pt}
`r round(stands %>% filter(stand == std) %>% .$veneer_vol/1000,1)` MBF veneer, `r round(stands %>% filter(stand == std) %>% .$saw_vol/1000,1)` MBF sawtimber, `r round(stands %>% filter(stand == std) %>% .$tie_vol/1000,1)` MBF tie logs,   
`r round(stands %>% filter(stand == std) %>% .$cord_vol)` cds pulp   

> - __Size class structure (%BA):__  	 
\vspace{2pt}
\indent 6-10”: `r round(sum(trees$stand == std & trees$dbh >= 6 & trees$dbh <= 10 & trees$live == 1)/sum(trees$stand == std & trees$dbh >= 6 & trees$live == 1)*100)`% | 11-16”: `r round(sum(trees$stand == std & trees$dbh > 10 & trees$dbh <=16 & trees$live == 1)/sum(trees$stand == std & trees$dbh >= 6 & trees$live == 1)*100)`%	| 17-22”: `r round(sum(trees$stand == std & trees$dbh > 16 & trees$dbh <= 22 & trees$live == 1)/sum(trees$stand == std & trees$dbh >= 6 & trees$live == 1)*100)`% |	23+”: `r round(sum(trees$stand == std & trees$dbh > 22 & trees$live == 1)/sum(trees$stand == std & trees$dbh >= 6 & trees$live == 1)*100)`%  

##Inventory information

> - `r stands %>% filter(stand == std) %>% .$plots` points, `r baf` BAF, `r paste(lubridate::month(month, label = TRUE, abbr = FALSE), year, sep = ", ")`

```{r stockingchart1, fig.cap="Points represent individual plots. Asterisk represnts stand average. Radial lines are quadratic stand diameters.", echo=FALSE, results='hide'}
this_stand <- stands %>%
  filter(stand == std)

# qsd lines --------------

dbhbold <- seq(2, 22, 2)
dbhlight <- seq(1, 21, 2)

# qsd labels -------------

dbhlabs <- paste(c(4,6,8,10,12,14,16,20), '"')

dbhlabstop <- tibble(y = rep(max(this_stand$ay1, 
                             max((plots %>% 
                                    filter(stand == std))$ba_live))+5, 
                             times = 8)) %>% 
  mutate(x = y/(.005454*c(4,6,8,10,12,14,16,20)^2))

dbhlabsloc <- tibble(x = c(493, 443, 372, 293, 216, 164, 127, 
                           82),
                     y = c(43, 87, 130, 160, 170, 175, 177, 
                           178.8))

# plot with a b & c curves -----------

curveplot <- plots %>% filter(stand == std) %>%
  ggplot(aes(tpa_live, ba_live)) +
  geom_abline(slope = .005454*dbhbold^2, col = "dark gray") +
  geom_abline(slope = .005454*dbhlight^2, col = "gray") +
  annotate("text", x = dbhlabsloc$x, 
           y = dbhlabsloc$y, 
           label = dbhlabs,
           color = "gray55",
           family = "Perpetua") +
  geom_curve(aes(x = this_stand$ax1,
                 y = this_stand$ay1,
                 xend = this_stand$ax2,
                 yend = this_stand$ay2),
             curvature = this_stand$acurv,
             angle = this_stand$aangle) + 
  geom_curve(aes(x = this_stand$bx1,
                 y = this_stand$by1,
                 xend = this_stand$bx2,
                 yend = this_stand$by2),
             curvature = this_stand$bcurv,
             angle = this_stand$bangle) + 
  geom_curve(aes(x = this_stand$cx1,
                 y = this_stand$cy1,
                 xend = this_stand$cx2,
                 yend = this_stand$cy2),
             curvature = this_stand$ccurv,
             angle = this_stand$cangle) +
  geom_text(aes(this_stand$ax2, 
                this_stand$ay2, 
                label = this_stand$alab),
            nudge_x = this_stand$anudge,
            family = "Perpetua") +
  geom_text(aes(this_stand$bx2, 
                this_stand$by2, 
                label = this_stand$blab),
            nudge_x = this_stand$bnudge,
            family = "Perpetua") +
  geom_text(aes(this_stand$cx2, 
                this_stand$cy2, 
                label = this_stand$clab),
            nudge_x = this_stand$cnudge,
            family = "Perpetua") +
  geom_point() +
  geom_point(aes(this_stand$tpa, 
                 this_stand$mean_ba),
             size = 3, 
             shape = 8, 
             stroke = 1.5) +
  theme(text = element_text(family = "Perpetua"), 
        legend.position = "right") +
  scale_x_continuous(name = "trees per acre") +
  scale_y_continuous(name = "basal area per acre (sq ft)") +
  labs(title = "Stocking chart",
       caption = paste("Reproduced from ", this_stand$name, 
                       " stocking guide: ", this_stand$author,
                       ". ", this_stand$source, sep = ""))

# plot w/o curves -----------------
  
curvelessplot <- plots %>% filter(stand == std) %>%
  ggplot(aes(tpa_live, ba_live)) +
  geom_abline(slope = .005454*dbhbold^2, col = "dark gray") +
  geom_abline(slope = .005454*dbhlight^2, col = "gray") +
  annotate("text", x = dbhlabsloc$x, 
           y = dbhlabsloc$y, 
           label = dbhlabs,
           color = "gray55",
           family = "Perpetua") +
  geom_point() +
  geom_point(aes(this_stand$tpa, 
                 this_stand$mean_ba),
             size = 3, 
             shape = 8, 
             stroke = 1.5) +
  theme(text = element_text(family = "Perpetua"), 
        legend.position = "right") +
  scale_x_continuous(name = "trees per acre") +
  scale_y_continuous(name = "basal area per acre (sq ft)") +
  labs(title = "Stocking chart")

ifelse(this_stand$structure.1[1] == "Even-aged", print(curveplot),
       print(curvelessplot))
```

```{r ags_ugs_table1, results="asis"}
temp1 <- stands %>% filter(stand == std)
temp2 <- c(temp1$mean_ba, temp1$qsd, temp1$tpa)
temp3 <- c(temp1$ba_ags, temp1$qsd_ags, temp1$tpa_ags)
temp4 <- c(temp1$ba_inv, temp1$qsd_in, temp1$tpa_inv)
temp0 <- c('Basal area (sqft/ac)', 'QSD (in)', 'Stems/ac')
temp5 <- data.frame(temp0, temp2, temp3, temp4)
names(temp5) <- c(' ', 'Total', 'Acceptable', 'Investment-grade')

kable(temp5, booktabs = T, digits = 0, caption = "Measures of stocking for all live trees (total), acceptable growing stock, and investment-grade growing stock (which is a subset of acceptable growing stock).")
```