#Area `r i` 
`r (stands %>% filter(stand == i))$type`    
\noindent `r sprintf("%.2f", (stands %>% filter(stand == i))$acres_legal)` legal acres | `r sprintf("%.2f", (stands %>% filter(stand == i))$acres_calc)` measured acres    

##Site-specific information

> - __Soils:__   
\indent\indent `r paste((soils %>% filter(stand == i))$text, collapse = " ") %>% str_replace_all("START", "\n") %>% str_replace_all("END", "  ")` 

> - __Site Class:__   
\vspace{2pt}
`r (stands %>% filter(stand == i))$site_class` (determined from soil mapping and field assessment)   

> - __Access:__   
\vspace{2pt}
`r (stands %>% filter(stand == i))$access` 

> - __Stand history:__   
\vspace{2pt}
`r (stands %>% filter(stand == i))$history` 

##Current forest information

> - __Age Class Structure:__   
\vspace{2pt}
`r (stands %>% filter(stand == i))$structure.1`  
```{r fig.margin=TRUE, fig.width=2, fig.height=6, fig.cap="Distributions are approximated with kernel density estimation. Common species are those that account for at least 8 percent of the total stocking and areas under each curve represent species basal areas."}

# diameter dist of main spp in a specific stand ----

trees %>% filter(stand == i, live == 1, dbh >= 6) %>%
  group_by(spp) %>%
  mutate(spp_sum = sum(live)) %>%
  ungroup() %>%
  mutate(spp = reorder(spp, -live, FUN = sum)) %>%
  filter(spp_sum/n()>=.08) %>% 
  ggplot(aes(dbh, y = ..count..)) +
  geom_density(alpha = .6, position = "identity", fill = "grey", col = "dark grey") +
  theme(text = element_text(family = "Perpetua"), 
        legend.position = "none") +
  facet_wrap(~spp, ncol = 1) +
  scale_y_continuous("relative basal area", labels = NULL, 
                     breaks = NULL) +
  scale_x_continuous("dbh") +
  labs(title = "Diameter distributions", 
       subtitle = "for common species")
```

```{r}
x <- trees %>% 
  filter(stand == i, live == 1, dbh >= 6) %>% 
  group_by(spp) %>% 
  summarize(spp_sum = sum(live)) %>%
  mutate(pct = spp_sum / sum(spp_sum)) %>%
  select(spp, pct) %>%
  arrange(-pct)
```

> - __Species (% stocking):__   
\vspace{2pt}
`r if_else(nrow(x)>0, paste(x$spp, " (", round(x$pct*100), '%)', sep = '', collapse = ", "), "none")`  

> - __Regeneration:__   
\vspace{2pt}
`r (stands %>% filter(stand == i))$regen`  

> - __Forest health:__   
\vspace{2pt}
`r (stands %>% filter(stand == i))$health`  
 

```{r}
if(any(trees$vigor == 5)){
  dead <- TRUE
  six <- round((sum(filter(trees, stand == i, dbh >= 6, dbh <= 10)$live == 0)*baf)/
                 (filter(stands, stand == i)$plots), 1)
  eleven <- round((sum(filter(trees, stand == i, dbh > 10, dbh <= 16)$live == 0)*baf)/
                    (filter(stands, stand == i)$plots), 1)
  seventeen <- round((sum(filter(trees, stand == i, dbh > 17, dbh <= 22)$live == 0)*baf)/
                       (filter(stands, stand == i)$plots), 1)
  twentythree <- round((sum(filter(trees, stand == i, dbh > 22)$live == 0)*baf)/
                         (filter(stands, stand == i)$plots), 1)
} else {
  dead <- FALSE
}
```

`r ifelse(dead, paste0('> - __Standing dead wood (sq ft/ac by size class):__   \n \\vspace{2pt} \\indent \\small 6-10": ', six, ' | 11-16”: ', eleven, ' | 17-22”: ', seventeen, ' | 23+”: ', twentythree), "")`

##Inventory information

> - `r stands %>% filter(stand == i) %>% .$plots` points, `r baf` BAF, `r paste(lubridate::month(month, label = TRUE, abbr = FALSE), year, sep = ", ")`

```{r fig.cap="Points represent individual plots. Asterisk represnts stand average. Radial lines are quadratic stand diameters.", echo=FALSE, results='hide'}
this_stand <- stands %>%
  filter(stand == i)

# qsd lines --------------

dbhbold <- seq(2, 26, 2)
dbhlight <- seq(1, 25, 2)

# qsd labels -------------

dbhlabs <- paste(c(4,6,8,10,12,14,16,20), '"', sep = "")

dbhlabstop <- tibble(y = rep(max(this_stand$ay1, 
                             max((plots %>% 
                                    filter(stand == i))$ba_live))+5, 
                             times = 8)) %>% 
  mutate(x = y/(.005454*c(4,6,8,10,12,14,16,20)^2))

dbhlabsright <- tibble(x = rep(max(this_stand$ax2,
                                   max((plots %>%
                                          filter(stand == i))$tpa_live))+50,
                               times = 8))
dbhlabsright <- dbhlabsright %>%
  mutate(y = x*.005454*c(4,6,8,10,12,14,16,20)^2)

# plot with a b & c curves -----------

curveplot <- plots %>% filter(stand == i) %>%
  ggplot(aes(tpa_live, ba_live)) +
  geom_abline(slope = .005454*dbhbold^2, col = "dark gray") +
  geom_abline(slope = .005454*dbhlight^2, col = "gray") +
  annotate("text", x = dbhlabstop$x, 
           y = dbhlabstop$y, 
           label = dbhlabs,
           color = "gray55",
           family = "Perpetua") +
  annotate("text", x = dbhlabsright$x, 
           y = dbhlabsright$y, 
           label = dbhlabs,
           color = "gray55",
           family = "Perpetua") +
  geom_curve(aes(x = this_stand$ax1,
                 y = this_stand$ay1,
                 xend = this_stand$ax2,
                 yend = this_stand$ay2),
             curvature = this_stand$acurv,
             angle = this_stand$aangle) + 
  geom_curve(aes(x = this_stand$bx1,
                 y = this_stand$by1,
                 xend = this_stand$bx2,
                 yend = this_stand$by2),
             curvature = this_stand$bcurv,
             angle = this_stand$bangle) + 
  geom_curve(aes(x = this_stand$cx1,
                 y = this_stand$cy1,
                 xend = this_stand$cx2,
                 yend = this_stand$cy2),
             curvature = this_stand$ccurv,
             angle = this_stand$cangle) +
  geom_text(aes(this_stand$ax2, 
                this_stand$ay2, 
                label = this_stand$alab),
            nudge_x = this_stand$anudge,
            family = "Perpetua") +
  geom_text(aes(this_stand$bx2, 
                this_stand$by2, 
                label = this_stand$blab),
            nudge_x = this_stand$bnudge,
            family = "Perpetua") +
  geom_text(aes(this_stand$cx2, 
                this_stand$cy2, 
                label = this_stand$clab),
            nudge_x = this_stand$cnudge,
            family = "Perpetua") +
  geom_point() +
  geom_point(aes(this_stand$tpa, 
                 this_stand$mean_ba),
             size = 3, 
             shape = 8, 
             stroke = 1.5) +
  theme(text = element_text(family = "Perpetua"), 
        legend.position = "right") +
  scale_x_continuous(name = "trees per acre",
                     limits = c(min(this_stand$cx1, 
                                    min((plots %>% 
                                           filter(stand == i))$tpa_live)),
                                max(this_stand$ax2,
                                    max((plots %>% 
                                           filter(stand == i))$tpa_live))+50)) +
  scale_y_continuous(name = "basal area per acre (sq ft)",
                     limits = c(min(this_stand$cy2, 
                                    min((plots %>% 
                                           filter(stand == i))$ba_live)),
                                max(this_stand$ay1,
                                    max((plots %>% 
                                           filter(stand == i))$ba_live))+5)) +
  labs(title = "Stocking chart",
       caption = paste("Reproduced from ", this_stand$name, 
                       " stocking guide: ", this_stand$author,
                       ". ", this_stand$source, sep = ""))

# plot w/o curves -----------------

dbhlabstop <- tibble(y = rep(max((plots %>% 
                                    filter(stand == i))$ba_live)+20, 
                             times = 8)) %>% 
  mutate(x = y/(.005454*c(4,6,8,10,12,14,16,20)^2))

dbhlabsright <- tibble(x = rep(max((plots %>%
                                          filter(stand == i))$tpa_live)+100,
                               times = 8)) %>%
  mutate(y = x*.005454*c(4,6,8,10,12,14,16,20)^2)
  
curvelessplot <- plots %>% filter(stand == i) %>%
  ggplot(aes(tpa_live, ba_live)) +
  geom_abline(slope = .005454*dbhbold^2, col = "dark gray") +
  geom_abline(slope = .005454*dbhlight^2, col = "gray") +
  annotate("text", x = dbhlabstop$x, 
           y = dbhlabstop$y, 
           label = dbhlabs,
           color = "gray55",
           family = "Perpetua") +
  annotate("text", x = dbhlabsright$x, 
           y = dbhlabsright$y, 
           label = dbhlabs,
           color = "gray55",
           family = "Perpetua") +
  geom_point() +
  geom_point(aes(this_stand$tpa, 
                 this_stand$mean_ba),
             size = 3, 
             shape = 8, 
             stroke = 1.5) +
  theme(text = element_text(family = "Perpetua"), 
        legend.position = "right") +
  scale_x_continuous(name = "trees per acre",
                     limits = c(min((plots %>% 
                                           filter(stand == i))$tpa_live)-100,
                                max((plots %>% 
                                           filter(stand == i))$tpa_live)+100)) +
  scale_y_continuous(name = "basal area per acre (sq ft)",
                     limits = c(min((plots %>% 
                                           filter(stand == i))$ba_live)-20,
                                max((plots %>% 
                                           filter(stand == i))$ba_live)+20)) +
  labs(title = "Stocking chart")

ifelse(this_stand$structure.1[1] == "Even-aged", print(curveplot),
       print(curvelessplot))
```

```{r results="asis"}
temp1 <- stands %>% filter(stand == i)
temp2 <- c(temp1$mean_ba, temp1$qsd, temp1$tpa)
temp3 <- c(temp1$ba_ags, temp1$qsd_ags, temp1$tpa_ags)
temp4 <- temp2 - temp3
temp4[2] <- sqrt((temp4[1]/temp4[3])/0.005454)
# temp4 <- c(temp1$ba_inv, temp1$qsd_in, temp1$tpa_inv)
temp0 <- c('Basal area (sq ft/ac)', 'QSD (in)', 'Stems/ac')
temp5 <- data.frame(temp0, temp2, temp3, temp4)

kable(temp5, booktabs = T, digits = 0, 
      col.names = c('Measure', 'Total', 'AGS', 'UGS'), 
      caption = "Measures of stocking for all live trees (Total), acceptable growing stock (AGS), and unacceptable growing stock (UGS).")
```

```{r results="asis"}
plts <- stands %>% filter(stand == i) %>% .$plots
resid <- ifelse(is.na(filter(stands, stand == i)$resid_ba_6_12_16_22), 0, 1)

temp1 <- trees %>% filter(stand == i, live == 1) %>% 
  mutate(size_class = cut(dbh, c(6, 12, 16, 22, Inf), 
                          include.lowest = T, 
                          right = F)) %>% 
  filter(!is.na(size_class)) %>% 
  group_by(size_class) %>% 
  summarize(total = n()*baf/plts,
            ags = sum(ags == 1)*baf/plts,
            ugs = total - ags) %>%
  mutate(size_class = c('6-11 in.', '12-15 in.', '16-21 in.', '22+ in.'))

header <- c("Size Class", "Total", "AGS", "UGS")
cap <- "Current basal area (sq ft/ac) of total growing stock, acceptable growing stock, and unacceptable growing stock by size class."

if (resid == 1) {
  resids <- as.list(str_split(filter(stands, stand == i)$resid_ba_6_12_16_22, "\\D+"))
  resids <- as.numeric(resids[[1]])
  
  temp1 <- temp1 %>% 
    mutate(target = resids)
  
  header <- c(header, "Target")
  cap <- "Current (total, acceptable, and unacceptable growing stock) and post-harvest target basal areas (sq ft/ac) by size class."
}

func <- function(z) if (is.numeric(z)) sum(z) else ''
sumrow <- as.data.frame(lapply(temp1, func), stringsAsFactors = F)
sumrow$size_class[1] <- "Total"
temp1 <- rbind(temp1, sumrow)

kable(temp1, booktabs = T, digits = 0, col.names = header, caption = cap)
```

##Long-term management system

**`r filter(stands, stand == i)$long_term_name`**

`r filter(stands, stand == i)$long_term`

##Silvicultural prescription

`r ifelse(!is.na(filter(stands, stand == i)$treatment_name), paste0("**", filter(stands, stand == i)$treatment_name, "**"), "")`    
`r ifelse(!is.na(filter(stands, stand == i)$treatment_yr), paste("\\noindent __Year:__", filter(stands, stand == i)$treatment_yr), "")`

`r filter(stands, stand == i)$treatment`

\newpage
